Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-29 paper/ijcai16.tex
paper: almost there, three lines missing

A straightforward encoding of the above desiderata leads to the
following mixed integer {\em non-linear} optimization problem over the
variables $\mu \in \bbR$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge \mu \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
    \\
        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
Let us illustrate the above piece by piece. The objective is composed
of four parts: we maximize the shared margin $\mu$ (first part) and
minimize the total sum of the ranking errors $\veps^i$ incurred by
each weight vector $\vw^{i}$ (second part), while at the same time
regularizing the magnitude of the weights (third part) and the quality
of the configurations $\{ \vx^{i} \}$ (last part). The non-negative
hyperparameters $\alpha,\beta,\gamma$ control the influence of the
various components. The weight regularization term copes with the
common scenario in which the user has strong preferences about
some attributes, but is indifferent to most of
them. %~\cite{} \stefano{@Paolo: add refs}.
The $\ell_1$ penalty is frequently used to improve the sparsity of
learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent
gains in generalization ability and efficiency, as confirmed by our
empirical findings (see Section~\ref{sec:experiments}).

The two core constraints are~(\ref{eq:wyconstr}) and~(\ref{eq:wxconstr}),
which are taken directly from the discussion above. The former enforces the
correct ranking of the observed user preferences, while the latter ensures that
the generated configurations are diverse in terms of the weight vectors they
maximize. Constraint~(\ref{eq:wbounds}), intended to be component-wise, and~(\ref{eq:xbounds})
ensure that the weights and configurations are feasible, while the
remaining two guarantee the non-negativity of the slack and of the margin.
Since we require $\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the weights to be non-negative.

Note that we are choosing the configurations $\{ \vx^i \}$ and the weight vectors $\{
\vw^i \}$ {\em simultaneously}.
Look at Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a pair ($k=2$). 
The constraint \ref{eq:wxconstr} is represented by a red line, that partitions the space of
feasible utility weights in two parts (in general, there will be $k$ subregions).
Since we maximize the margin $\mu$, the optimizer
will prefer a set of configurations $\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space
in an even way.
In each subregion, we have corresponding utility weight vector $\vw^{i}$
lying ``close'' to the centre of each subregion. 
If we then present $\vx^{1}$ and $\vx^{2}$ to the user, and the user states that he prefers
the former to the latter, the feasible region for the weights will then become the part of the polytope to the left of the red line; moreover $\vw^{1}$ will maximize the margin in the classic sense ($k=1$).
%Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner.
%Therefore, setwise maxmargin can be thought as the set that, given any possible preference about the items, ensure that the reduction of utility uncertainty is the greatest
%\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}

@@

Substituting the above MILP constraints into the original non-linear
formulation, we obtain the following mixed-integer linear problem:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu
    \\
    %     & \;\; \forall \; i \in [k], \forall \; z \in [m] \nonumber
    % \\
        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [m] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \}
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \quad \veps^{i}_h \ge 0 \nonumber
    \\
        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
which can be solved by any suitable MILP solver.

@@

\begin{algorithm}
{\footnotesize
\begin{algorithmic}[1]
    \Procedure{SetMargin}{$k, \alpha, \beta, \gamma, T$}
        \State $\calD \gets \emptyset$
        \For{$t = 1, \ldots, T$}
            \State \{$\vw^{i}, \vx^{i}\}_{i=1}^k \gets \text{{\sc Solve}}(\calD, k, \alpha, \beta, \gamma)$
            \For{$\vx^{i},\vx^{j} \in \{ \vx^{1}, \ldots, \vx^{k} \} \; \text{{\bf s.t.}} \; i < j$}
                \State $\calD \gets \calD \cup \text{{\sc QueryUser}}(\vx^{i},\vx^{j})$
            \EndFor
        \EndFor
        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(\calD, 1, \alpha, \beta, \gamma)$
        \State ${\bf return}\; \vw^*, \vx^*$
    \EndProcedure
\end{algorithmic}
}
\caption{\label{alg:setmargin} The {\sc SetMargin} algorithm. Here $k$ is the
set size, $\alpha,\beta,\gamma$ are the hyperparameters, and $T$ is the maximum
number of iterations. The values of $\calX_\text{feasible}$, $\vw^\top$ and
$\vw^\bot$ are left implicit.}
\end{algorithm}

\paragraph{Linearly dependent real attributes.} In many domains of interest,
items are composed of both Boolean and real-valued attributes, where the latter
depend linearly  on the former. This is for instance the case for the price,
weight and power consumption of a laptop model, which depend linearly on the
choice of components.
%
In this setting, item configurations are composed of two parts: $\vx =
(\vx_B;\vx_R)$, where $\vx_B$ is Boolean and $\vx_R$ is real-valued, and the
relation between the two is entirely described by an appropriately sized
non-negative cost matrix $C$, such that $\vx_R = C \vx_B$. It is
straightforward to extend the MILP formulation to this setting, as follows.

First, we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
becomes:
%
$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C' \vw_R, \vx_B \rangle $$
%
The generalized problem is obtained by substituting $\vw^i$ with $\vv^i :=
\vw_B^i + C' \vw_R^i$.  All constraints remain the same. The only notable
change occurs in Eq.~(\ref{eq:wbounds2}), which becomes:
%
$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C' \vw_R^\bot) \le \vv^i \le (\vw_B^\top + C' \vw_R^\top)$$
%
As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
to be non-negative.

@@

\begin{figure*}[t]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        % \\
        % \hline \hline
        % \multicolumn{4}{c}{{\sc Uniform}}
        \\
        {\sc Uniform}  &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
         % \multicolumn{4}{c}{{\sc Normal}}
        \\
         {\sc Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        {\sc Sparse Uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        {\sc Sparse Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
      represents a different sampling distribution for user utility.
      The number of iterations is plotted against the utility loss
      (first and third columns) and the cumulative time (second and
      fourth columns). Thick lines indicate median values over users,
      while standard deviations are shown as shaded areas.}
\end{figure*}

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions and datasets $r=3$ and $r=4$. Solution quality is
measured in terms of utility loss $ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right)$,
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that {\sc riVOI},  {\sc QI} and
{\sc EUS} are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
%In all figures, thick lines indicate median values over the different
%users, shaded areas indicate standard deviations. Our {\sc SetMargin}
%algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
%in green and {\sc EUS} in gray. 
For all algorithms, one iteration corresponds to a single pairwise (we
used {\sc SetMargin} with $k=2$).
%query to the user 
%Results for $r=3$ and $r=4$ are reported.  
For dense weight vector distributions (first two rows), our approach
achieves results which are indistinguishable from the competitors in a
fraction of their time.  Indeed, all Bayesian approaches%
%\footnote{Note
%  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
%  inference greatly dominates the time spent for choosing the next
%  query to ask.\andrea{@paolo: necessario?}} 
 become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weights, 
while none of the other approaches is %explicitly 
designed to favour sparse functions.

@@

\begin{figure*}[t]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % \multicolumn{2}{c}{{\sc synthetic 4}} &
        % \multicolumn{2}{c}{{\sc synthetic 5}}
        % \\
        % {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
        % \\
        % \hline \hline
        {\sc uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
        \\
        \hline
        {\sc sparse normal} &        
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
        \\
        \hline
     \end{tabular}
    }
    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
      with $k=2,3,4$, in orange, blue and green respectively for the
      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions. Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
\end{figure*}

@@

A PC configurations is defined by eight attributes: computer type (laptop,
desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
size, and price.
%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by the sum of the price of its components plus a bias due to
branding.
%
Interactions between attributes are expressed as Horn clauses. An example
constraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4''%
%\footnote{Despite the components
%included in the dataset, which are a bit outdated, the dataset itself is pretty
%realistic.} \andrea{necessario?}
, which can be encoded into MILP form as
%
$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
%
Note that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1 because
of the mutual exclusivity constraints of the one-hot encoding.  The
dataset includes constraints between the following attributes:
manufacturer $\to$ type, manufacturer $\to$ CPU, type $\to$ RAM
amount, type $\to$ storage size, type $\to$ monitor size, for 16 Horn
constraints total. We do not report the full list due to space
limitations. Note that the search space is of the order of hundreds of
thousands of candidate configurations, and is far beyond reach of
existing Bayesian approaches.

>>>>

A straightforward encoding of the above desiderata leads to the
following mixed integer {\em non-linear} optimization problem over the
variables $\mu \in \bbR$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge \mu \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
%    \\
%        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
Let us illustrate the above piece by piece. The objective is composed
of four parts: we maximize the shared margin $\mu$ (first part) and
minimize the total sum of the ranking errors $\veps^i$ incurred by
each weight vector $\vw^{i}$ (second part), while at the same time
regularizing the magnitude of the weights (third part) and the quality
of the configurations $\{ \vx^{i} \}$ (last part). The non-negative
hyperparameters $\alpha,\beta,\gamma$ control the influence of the
various components. The weight regularization term copes with the
common scenario in which the user has strong preferences about some
attributes, but is indifferent to most of
them. %~\cite{} \stefano{@Paolo: add refs}.
The $\ell_1$ penalty is frequently used to improve the sparsity of
learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent
gains in generalization ability and efficiency, as confirmed by our
empirical findings (see
Section~\ref{sec:experiments}). Constraint~(\ref{eq:wyconstr})
enforces the correct ranking of the observed user preferences, while
constraint~(\ref{eq:wxconstr}) ensures that the generated
configurations are diverse in terms of the weight vectors they
maximize. Constraints~(\ref{eq:wbounds}) and~(\ref{eq:xbounds}) ensure
that the weights and configurations are feasible, while the last one
guarantees the non-negativity of the slacks.  Since we require
$\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the
weights to be non-negative.

Note that we are choosing the configurations $\{ \vx^i \}$ and the
weight vectors $\{ \vw^i \}$ {\em simultaneously}.  Look at
Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a
pair ($k=2$).  The constraint \ref{eq:wxconstr} is represented by a
red line, that partitions the space of feasible utility weights in two
parts (in general, there will be $k$ subregions).  Since we maximize
the margin $\mu$, the optimizer will prefer a set of configurations
$\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space in an even
way.  In each subregion, we have corresponding utility weight vector
$\vw^{i}$ lying ``close'' to its centre. If the user indicates a
preference for $\vx^{1}$ over $\vx^{2}$, the feasible region for the
weights will then become the part of the polytope to the left of the
red line; moreover $\vw^{1}$ will maximize the margin in the classic
sense ($k=1$).
%Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner.
%Therefore, setwise maxmargin can be thought as the set that, given any possible preference about the items, ensure that the reduction of utility uncertainty is the greatest
%\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}

@@

Substituting the above MILP constraints into the original non-linear
formulation, we obtain the following mixed-integer linear problem:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu
    \\
    %     & \;\; \forall \; i \in [k], \forall \; z \in [m] \nonumber
    % \\
        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [m] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \}
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \quad \veps^{i}_h \ge 0 \nonumber
%    \\
%        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
%which can be solved by any suitable MILP solver.

@@

\begin{algorithm}[t]
{\footnotesize
\begin{algorithmic}[1]
    \Procedure{SetMargin}{$k, \alpha, \beta, \gamma, T$}
        \State $\calD \gets \emptyset$
        \For{$t = 1, \ldots, T$}
            \State \{$\vw^{i}, \vx^{i}\}_{i=1}^k \gets \text{{\sc Solve}}(\calD, k, \alpha, \beta, \gamma)$
            \For{$\vx^{i},\vx^{j} \in \{ \vx^{1}, \ldots, \vx^{k} \} \; \text{{\bf s.t.}} \; i < j$}
                \State $\calD \gets \calD \cup \text{{\sc QueryUser}}(\vx^{i},\vx^{j})$
            \EndFor
        \EndFor
        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(\calD, 1, \alpha, \beta, \gamma)$
        \State ${\bf return}\; \vw^*, \vx^*$
    \EndProcedure
\end{algorithmic}
}
\caption{\label{alg:setmargin} The {\sc SetMargin} algorithm. Here $k$ is the
set size, $\alpha,\beta,\gamma$ are the hyperparameters, and $T$ is the maximum
number of iterations. The values of $\calX_\text{feasible}$, $\vw^\top$ and
$\vw^\bot$ are left implicit.}
\end{algorithm}

\paragraph{Linearly dependent real attributes.} In many domains of
interest, items are composed of both Boolean and real-valued
attributes, where the latter depend linearly on the former. This is
for instance the case for the price, weight and power consumption of a
laptop model, which depend linearly on the choice of components.  In
this setting, item configurations are composed of two parts:
$\vx = (\vx_B;\vx_R)$, where $\vx_B$ is Boolean and $\vx_R$ is
real-valued and can be written as $\vx_R = C \vx_B$ for an
appropriately sized non-negative cost matrix $C$.  It is
straightforward to extend the MILP formulation to this setting. First,
we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
becomes:
%
$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C' \vw_R, \vx_B \rangle $$
%
The generalized problem is obtained by substituting $\vw^i$ with $\vv^i :=
\vw_B^i + C' \vw_R^i$.  All constraints remain the same. The only notable
change occurs in Eq.~(\ref{eq:wbounds2}), which becomes:
%
$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C' \vw_R^\bot) \le \vv^i \le (\vw_B^\top + C' \vw_R^\top)$$
%
%As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
%to be non-negative.

@@

\begin{figure*}[t]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        % \\
        % \hline \hline
        % \multicolumn{4}{c}{{\sc Uniform}}
        \\
        {\sc Uniform}  &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
         % \multicolumn{4}{c}{{\sc Normal}}
        \\
         {\sc Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        {\sc Sparse Uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        {\sc Sparse Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        %\hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
      represents a different sampling distribution for user utility.
      The number of iterations is plotted against the utility loss
      (first and third columns) and the cumulative time (second and
      fourth columns). Thick lines indicate median values over users,
      while standard deviations are shown as shaded areas.}
\end{figure*}

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions and datasets $r=3$ and $r=4$. Solution quality is
measured in terms of utility loss $ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right)$,
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that {\sc riVOI},  {\sc QI} and
{\sc EUS} are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
%In all figures, thick lines indicate median values over the different
%users, shaded areas indicate standard deviations. Our {\sc SetMargin}
%algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
%in green and {\sc EUS} in gray. 
For all algorithms, one iteration corresponds to a single pairwise query (we
used {\sc SetMargin} with $k=2$).
%query to the user 
%Results for $r=3$ and $r=4$ are reported.  
For dense weight vector distributions (first two rows), our approach
achieves results which are indistinguishable from the competitors in a
fraction of their time.  Indeed, all Bayesian approaches
%\footnote{Note
%  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
%  inference greatly dominates the time spent for choosing the next
%  query to ask.\andrea{@paolo: necessario?}} 
 become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weights, 
while none of the other approaches is %explicitly 
designed to favour sparse functions.

@@

\begin{figure*}[t]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % \multicolumn{2}{c}{{\sc synthetic 4}} &
        % \multicolumn{2}{c}{{\sc synthetic 5}}
        % \\
        % {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
        % \\
        % \hline \hline
        {\sc uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
        \\
        \hline
        {\sc sparse normal} &        
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
        \\
        %\hline
     \end{tabular}
    }
    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
      with $k=2,3,4$, in orange, blue and green respectively for the
      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions. Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
\end{figure*}

@@

A PC configurations is defined by eight attributes: computer type (laptop,
desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
size, and price.
%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by the sum of the price of its components plus a bias due to
branding.
%
Interactions between attributes are expressed as Horn clauses (e.g. a certain manufacturer implies a set of possible CPUs).
% . An example
% constraint might look like this: ``if the manufacturer is Apple, then the
% CPU must be either a PowerPC G3 or a PowerPC G4''%
% %\footnote{Despite the components
% %included in the dataset, which are a bit outdated, the dataset itself is pretty
% %realistic.} \andrea{necessario?}
% , which can be encoded into MILP form as
% %
% $ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
% %
% Note that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1 because
% of the mutual exclusivity constraints of the one-hot encoding.  
The
dataset includes 16 Horn constraints (the full list is omitted for
space limitations).
% constraints between the following attributes:
% manufacturer $\to$ type, manufacturer $\to$ CPU, type $\to$ RAM
% amount, type $\to$ storage size, type $\to$ monitor size, for 16 Horn
% constraints total. We do not report the full list due to space
% limitations. 
Note that the search space is of the order of hundreds of thousands of
candidate configurations, and is far beyond reach of existing Bayesian
approaches.
